---
title: "Metadata"
author: "Wells"
date: "2026-09-15"
output: html_document
---

Class exercise to import your project metadata from the spreadsheet into R 

--Create a folder called "class_project" in your code folder 

--Save this file to the "class_project" folder 

--Move your Nexis spreadsheet to the folder 

--Create a project file on R Studio: File \| New Project \| \*\*\* select Existing folder \| "class_project" 

--After project.Rproj launches, then launch this file, metadata.Rmd, and complete the tutorial

**By doing this, you avoid the ring of hell where R can't find your data**

#load tidyverse, janitor, rio

```{r}

library(tidyverse)
library(janitor)
library(rio)
#install.packages("readxl")
library(readxl)

```

#Import your data

```{r}

data <- read_excel("Xinjiang_women.XLSX") |>
  clean_names()
View(data)

names(data)
```

# Clean names

```{r}
data <- data |> 
  clean_names()

```

# check data types
```{r}
glimpse(data)
```


Next Steps: Count years, and create a ggplot

# Transform date into years. See tutorial
```{r}

data <- data |> 
  mutate(year = year(published_date))

```

#count
```{r}
data |> 
  count(year)
```
year      n
2015      1
2018      1
2020      1
2021     62
2022      9
2023      9
2024      8
2025      5
2026      4

#publication
```{r}
data |> 
  count(publication_4)|>
  arrange(desc(n))
```

Top 10 publications and the number of articles

publication_4 - n

CE Noticias Financieras English	- 14			
China Daily	- 13			
The Independent (United Kingdom) -	13			
Global Times (China)	- 9			
China Daily European Edition	- 6			
China Daily (Hong Kong Edition)	- 5			
Financial Times (London, England)	- 4			
MailOnline	- 4			
Postmedia Breaking News	- 3			
The Times (London)	- 3	

#country
```{r}
data |> 
  count(countries)|>
arrange(desc(n))
```
Top 10 (plus NA) countries and the number of publications: 

countries - n

China, People's Republic of	- 30	
United Kingdom of Great Britain and Northern Ireland	- 18	
England & Wales	- 14	
Mexico	- 14	
Australia	- 5	
Canada	- 5	
China, People's Republic of\nHong Kong Special Administrative Region of China	- 4	
India, Republic of	- 3	
NA	- 3	
Botswana, Republic of	- 1	

#company
```{r}
data |> 
  count(company)|>
arrange(desc(n))
```
Top 8 companies and the number of publications:

company (including NA) - n

NA	- 70			
TWITTER INC	- 11			
ASSOCIATED PRESS	- 7			
FACEBOOK INC	- 4			
METROPOLITAN BANK HOLDING CORP	- 3			
NATIONAL PRESS CLUB	- 2			
EUROPEAN PARLIAMENT	- 1			
ROOTS CORP	- 1			
SEA LTD	- 1	

#length

```{r}

data_big <- data |> 
  mutate(length1=as.numeric(length)) |>
  filter(length1 > 2903 & length1 <2505)

  #count(length)|>
#arrange(desc(n))
```


# Create a ggplot
```{r}
data |> 
  count(year) |> 
   ggplot(aes(x = year, y = n,
             fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  scale_x_continuous(breaks=c(2015:2026)) + #fixes wonky x axis
  labs(title = "Years for Xinjiang Women Research",
       subtitle = "Limited NexisUni Run",
       caption = "Graphic by Aya, 9-15-2026",
       y="Number of Articles",
       x="")



```

#Count publications

```{r}
data |> 
  count(publication_4) |>
arrange(desc(n))
```
publication_4 - n

CE Noticias Financieras English	- 14			
China Daily	- 13			
The Independent (United Kingdom) - 13			
Global Times (China)	- 9			
China Daily European Edition	- 6			
China Daily (Hong Kong Edition)	- 5			
Financial Times (London, England)	- 4			
MailOnline	- 4			
Postmedia Breaking News	- 3			
The Times (London)	- 3	

#Visualize publications

```{r}
data |> 
  count(publication_4) |>
#arrange(desc(n)) |>
# ggplot(aes(x = publication_4, y = n,
            # fill = n)) +
  ggplot(aes(x=reorder(publication_4, n), y=n, fill=n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
 #scale_x_continuous(breaks=c(2015:2026)) + #fixes wonky x axis
  coord_flip() +
  labs(title = "Years for Xinjiang Women Research",
       subtitle = "Limited NexisUni Run",
       caption = "Graphic by Aya, 9-15-2026",
       y="Publication",
       x="")

```




# NOTES ON FIXING COMPLEX Date COLUMN WITH HOURS AND MINUTES

```{r}
#women <- women |> 
 # mutate(date = as.Date(strptime(published_date, format = "%B %d, %Y %A")))
```